Evaluation of Diagnostic Concordance Between Algorithms for Parkinson’s Disease Dementia

### 2025-06-25 _manuscript/manuscript.qmd
Put manuscript rendering directly inside the _targets script.

upstore <- here::here(tar_config_get("store"))
tar_source(here::here("R"))
data <- tar_read(raw_data, store = upstore)
crit <- tar_read(pdd_data, store = upstore)$criteria
vars <- tar_read(variables, store = upstore)

```

@@

```{r}
#| label: crit-check

critnum <-
  sapply(
    c("mmse", "moca", "smoca", "lvlII"),
    function(i)
      subset(crit, group == i) |>
      nrow()
  )

@@

```{r}
#| label: tbl-crits
#| tbl-cap: "Summary of probable PDD operationalizations compared in the study."

@@

```{r}
#| label: results

desc <- tar_read(sample_description, store = upstore)
prev <- tar_read(prevalence_summaries, store = upstore)
pred <- tar_read(demographic_predictors, store = upstore)
conc <- tar_read(concordance_statistics, store = upstore)
perc <- prev$table$perc

@@

# In-text rate estimates
rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", prev$table$IADL)], 2, f))
rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[prev$table$IADL == "FAQ > 7"], 2, f))

# Minimum p-value of logistic regressions
min_p <- subset(pred$values, quantity == "p-value")$estimate |> min() |> do_summary(3, "p")
```

@@

```{r}
#| label: tbl-desc
#| tbl-cap: "Cognitive characteristics of the sample."

@@

```{r}
#| label: fig-rates
#| fig-cap: Summary of the estimates of probable PDD rate.
#| apa-note: Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD.

prev$plot
```

@@

```{r}
#| label: kappas

algos <- list(
  faq_tot = subset(prev$table, IADL == "FAQ > 7")$type,
  faq_9 = subset(prev$table, IADL == "FAQ 9 > 1")$type
)

# Check for overlap:
continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)

# Prepare combinations within IADL operationalisation:
kappas <-
  lapply(
    set_names(names(algos)),
    function(o)
      combn(algos[[o]], 2) |>
      t() |>
      as_tibble() |>
      `colnames<-`(c("predictor", "reference")) |>
      left_join(conc$table, by = c("predictor", "reference")) |>
      select(Kappa_raw) |>
      pull()
  )

@@

```{r}
#| label: fig-acc
#| fig-cap: Prediction accuracy matrix.
#| apa-note: The matrix depicts classification accuracy of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis. Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response. Cases with asterisk indicate predictive accuracy statistically significantly higher than the No Information Rate.

@@

```{r}
#| label: lvlII-stats

lvlII_rates <- c(
  faq_tot = subset(prev$table, type == "Lvl.II (1)")$perc |> do_summary(2) |> paste0("%"),
  faq_9 = subset(prev$table, type == "Lvl.II (2)")$perc |> do_summary(2) |> paste0("%")
)

@@

```{r}
#| label: tbl-approx
#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference."

@@

```{r}
#| label: rates-minmax

@@

```{r}
#| label: tbl-algos
#| tbl-cap: Summary of all algortihms for probable PDD used in the study.

prev$gtables$algorithms
```

@@

```{r}
#| label: tbl-estimands
#| tbl-cap: Mapping between research objectives and quantities of interest in the current study.

@@

```{r}
#| label: tbl-rates
#| tbl-cap: Estimates of the rate of probable PDD in the sample.

prev$gtables$rates
```

```{r}
#| label: tbl-approx-bal
#| tbl-cap: Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score.

@@

```{r}
#| label: fig-data
#| fig-cap: Representation of study data.
#| apa-note: The figure shows whether patients (x-axis) ordered from the youngest (left) to the oldest (right) were classified as probable PDD by each tested algorithm (y-axis) ordered from the one with the lowest (bottom) to the highest (top) PDD rate estimate. Patients printed in red are women, patients printed in blue are men. Red cells indicate probable PDD diagnosis, grey cells indicate non-PDD diagnosis and white cells indicate missing diagnosis.

@@

```{r}
#| label: fig-pars
#| fig-cap: Summary of logistic regressions parameters prediction probable PDD by age and sex.
#| apa-note: Histograms represent odd ratio (OR) estimates and p-values associated with age, sex, and their interaction as predictors of each of the 68 probable PDD classification. In the case of parameters for sex, values higher than 20 were omitted for clarity. Vertical lines indicate OR = 1 and p = .05.

@@

```{r}
#| label: fig-kappa
#| fig-cap: Cohen's κ matrix.
#| apa-note: The matrix depicts Cohen's κ measuring agreement between algorithms for PDD. Algorithms printed in red defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.

@@

```{r}
#| label: fig-sens
#| fig-cap: Sensitivity matrix.
#| apa-note: The matrix depicts sensitivity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis. Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.

@@

```{r}
#| label: fig-spec
#| fig-cap: Specificity matrix.
#| apa-note: The matrix depicts specificity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis. Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response.

>>>>

upstore <- here::here(tar_config_get("store"))
tar_source(here::here("R"))
data <- tar_read(raw_data, store = upstore)
crit <- tar_read(pdd_data, store = upstore)$criteria
vars <- tar_read(variables, store = upstore)
```

@@

```{r}
#| label: "crit-check"

critnum <- sapply(
  c("mmse", "moca", "smoca", "lvlII"), function(i) {
    subset(crit, group == i) |>
      nrow()
  }
)

@@

```{r}
#| label: "tbl-crits"
#| tbl-cap: "Summary of probable PDD operationalizations compared in the study."

@@

```{r}
#| label: "results"

desc <- tar_read(sample_description, store = upstore)
rates <- tar_read(rate_summaries, store = upstore)
pred <- tar_read(demographic_predictors, store = upstore)
conc <- tar_read(concordance_statistics, store = upstore)
perc <- rates$table$perc

@@

# In-text rate estimates
rates_overall <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc, 2, f))
rates_FAQ9 <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[grepl("FAQ 9", rates$table$IADL)], 2, f))
rates_FAQtot <- sapply(c("M", "SD", "Md", "minmax"), function(f) do_summary(perc[rates$table$IADL == "FAQ > 7"], 2, f))

# Minimum p-value of logistic regressions:
min_p <-
  subset(pred$values, quantity == "p-value")$estimate |>
  min() |>
  do_summary(3, "p")

# List of algorithms:
algos <- list(
  faq_tot = subset(rates$table, IADL == "FAQ > 7")$type,
  faq_9 = subset(rates$table, IADL == "FAQ 9 > 1")$type
)

continue <- !(any(algos$faq_tot %in% algos$faq_9) || any(algos$faq_9 %in% algos$faq_tot))
stopifnot("Some algorithm(s) use both IADL operationalisations!" = continue)
```

@@

```{r}
#| label: "tbl-desc"
#| tbl-cap: "Cognitive characteristics of the sample."

@@

```{r}
#| label: "fig-rates"
#| fig-cap: "Summary of the estimates of probable PDD rate."
#| apa-note: "Vertical lines represent estimates arrived at by using sMoCA (dotted) or Level II (dashed) with FAQ item 9 (orange) or FAQ total score (blue) as criteria for probable PDD."

rates$plot
```

@@

```{r}
#| label: "kappas"

# Prepare combinations within IADL operationalisation:
kappas <- lapply(
  set_names(names(algos)), function(o) {
    combn(algos[[o]], 2) |>
      t() |>
      as_tibble() |>
      `colnames<-`(c("predictor", "reference")) |>
      left_join(conc$table, by = c("predictor", "reference")) |>
      select(Kappa_raw) |>
      pull()
  }
)

@@

```{r}
#| label: "fig-acc"
#| fig-cap: "Prediction accuracy matrix."
#| apa-note: "The matrix depicts classification accuracy of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis. Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response. Cases with asterisk indicate predictive accuracy statistically significantly higher than the No Information Rate."

@@

```{r}
#| label: "lvlII-stats"

lvlII_rates <- c(
  faq_tot = subset(rates$table, type == "Lvl.II (1)")$perc |> do_summary(2) |> paste0("%"),
  faq_9 = subset(rates$table, type == "Lvl.II (2)")$perc |> do_summary(2) |> paste0("%")
)

@@

```{r}
#| label: "tbl-approx"
#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference."

@@

```{r}
#| label: "rates-minmax"

@@

```{r}
#| label: "tbl-algos"
#| tbl-cap: "Summary of all algortihms for probable PDD used in the study."

rates$gtables$algorithms
```

@@

```{r}
#| label: "tbl-estimands"
#| tbl-cap: "Mapping between research objectives and quantities of interest in the current study."

@@

```{r}
#| label: "tbl-rates"
#| tbl-cap: "Estimates of the rate of probable PDD in the sample."

rates$gtables$rates
```

```{r}
#| label: "tbl-approx-bal"
#| tbl-cap: "Level I algorithms for probable PDD as predictors of Level II classification as the reference arranged by their balanced accuracy score."

@@

```{r}
#| label: "fig-data"
#| fig-cap: "Representation of study data."
#| apa-note: "The figure shows whether patients (x-axis) ordered from the youngest (left) to the oldest (right) were classified as probable PDD by each tested algorithm (y-axis) ordered from the one with the lowest (bottom) to the highest (top) PDD rate estimate. Patients printed in red are women, patients printed in blue are men. Red cells indicate probable PDD diagnosis, grey cells indicate non-PDD diagnosis and white cells indicate missing diagnosis."

@@

```{r}
#| label: "fig-pars"
#| fig-cap: "Summary of logistic regressions parameters prediction probable PDD by age and sex."
#| apa-note: "Histograms represent odd ratio (OR) estimates and p-values associated with age, sex, and their interaction as predictors of each of the 68 probable PDD classification. In the case of parameters for sex, values higher than 20 were omitted for clarity. Vertical lines indicate OR = 1 and p = .05."

@@

```{r}
#| label: "fig-kappa"
#| fig-cap: "Cohen's κ matrix."
#| apa-note: "The matrix depicts Cohen's κ measuring agreement between algorithms for PDD. Algorithms printed in red defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response."

@@

```{r}
#| label: "fig-sens"
#| fig-cap: "Sensitivity matrix."
#| apa-note: "The matrix depicts sensitivity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis. Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response."

@@

```{r}
#| label: "fig-spec"
#| fig-cap: "Specificity matrix."
#| apa-note: "The matrix depicts specificity of algorithms for PDD depicted on x-axis in predicting outcomes based on algorithms on the y-axis. Algorithms printed in blue defined IADL deficit by FAQ total score, algorithms printed in black defined IADL deficit by FAQ item 9 response."
